---
title: "ANLY 545 Project"
output: html_document
Author: "Gun Violence in the US"
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Load all required packages
library(tidyverse); library(leaflet); library(rgdal); library(sp); library(readr); library(plotly); library(scales);
library(kableExtra); library(treemapify); library(expss)

#load the data - set working directory to the data folder first
gunviolence <- read.csv("gun-violence-data.csv")
```

## Introduction

```{r setup, include=FALSE, warning=FALSE, message=FALSE}

```

## Data Source

```{r setup, include=FALSE, warning=FALSE, message=FALSE}

```

## Methodology

```{r setup, include=FALSE, warning=FALSE, message=FALSE}

```

## Analysis and Visualization

```{r setup, include=FALSE, warning=FALSE, message=FALSE}

```
###Gun incident over year

library(ggplot2)

p<-gun_violence%>%
  ggplot(aes(x=as.factor(year))) + geom_bar(stat='count', fill = 'darkgreen') +
  scale_y_continuous(labels=comma) +
  geom_label(stat = "count",aes(label = ..count.., y = ..count..))

p + ggtitle("Gun Incident Over Year") +
  xlab("Year") + ylab("Number of gun incidents")
```


## Conclusion

```{r setup, include=FALSE, warning=FALSE, message=FALSE}

```
